---
title: "Divy Case Study"
author: "Khalid Bouskou"
date: "11/19/2021"
---

## Load Library
This section to load library i need for my case study .

```{r}
library(tidyverse)
library(lubridate)
```
## Load Raw Dataset
This chunk for loading all data file in csv i need .

```{r}
M2020_11 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2020_11.csv")
M2020_12 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2020_12.csv")
M2021_01 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_01.csv")
M2021_02 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_02.csv")
M2021_03 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_03.csv")
M2021_04 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_04.csv")
M2021_05 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_05.csv")
M2021_06 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_06.csv")
M2021_07 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_07.csv")
M2021_08 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_08.csv")
M2021_09 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_09.csv")
M2021_10 <- read.csv("C:/Users/KasiolEr/Desktop/CSV/M2021_10.csv")
```
## Checking Dataset
This chunk for checking data types and names of variables and dimensions of every dataset .
```{r echo=FALSE}
str(M2020_11)
str(M2020_12)
str(M2021_01)
str(M2021_02)
str(M2021_03)
str(M2021_04)
str(M2021_05)
str(M2021_06)
str(M2021_07)
str(M2021_08)
str(M2021_09)
str(M2021_10)
```


## Convert data type 
after checking each dataset I find that the first dataset M2020_11
have two variable doesn't match two other dataset left 
and this two variable are 'start_station_id' and 'end_station_id' by default they are in numeric so i run this chunk to convert it to charachter for mergin it in the next step .
```{r}
M2020_11 <- mutate(M2020_11, start_station_id = as.character(start_station_id))
M2020_11 <- mutate(M2020_11, end_station_id = as.character(end_station_id))

```


## Combine 3 dataset into the 1 quarter 
I use R to combine dataset monthly into quarter year
My machine excel software can't handle merging many dataset into one .

```{r}
##
Q1 <- bind_rows(M2020_11, M2020_12, M2021_01)
Q2 <- bind_rows(M2021_02, M2021_03, M2021_04)
Q3 <- bind_rows(M2021_05, M2021_06, M2021_07)
Q4 <- bind_rows(M2021_08, M2021_09, M2021_10)

## Export quarter dataset as csv file .
#Note : row.names parameter in False they not add a new column for names as 'X' .

write.csv(Q1, "C:/Users/KasiolEr/Desktop/Q1.csv", row.names = FALSE)
write.csv(Q2, "C:/Users/KasiolEr/Desktop/Q2.csv", row.names = FALSE)
write.csv(Q2, "C:/Users/KasiolEr/Desktop/Q2.csv", row.names = FALSE)
write.csv(Q2, "C:/Users/KasiolEr/Desktop/Q2.csv", row.names = FALSE)

```



## Load Dataset
This section to load data from local machine
Note : Make sure you change the name of your machine before load the data 
```{r}
Q1 <- read.csv("C:/Users/KasiolEr/Desktop/Q1.csv")
Q2 <- read.csv("C:/Users/KasiolEr/Desktop/Q2.csv")
Q3 <- read.csv("C:/Users/KasiolEr/Desktop/Q3.csv")
Q4 <- read.csv("C:/Users/KasiolEr/Desktop/Q4.csv")

## This code to merge all quarter period into a year .
annual_data <- bind_rows(Q1, Q2, Q3, Q4)

## Export annual dataset as csv file .
write.csv(annual_data, "C:/Users/KasiolEr/Desktop/annual.csv", row.names = FALSE)
```

## Check the new dataset 

```{r}
## This chunk for checking our new dataset consistency and interoperability .
table(Q1$rideable_type, Q1$member_casual)
table(Q2$rideable_type, Q2$member_casual)
table(Q3$rideable_type, Q3$member_casual)
table(Q4$rideable_type, Q4$member_casual)
table(annual_data$rideable_type, annual_data$member_casual)

## This code to making a copy of annual data  for further analysis .
annual_v2 <- annual_data
```



## Descriptive analysis
This section for processing calculation and analysis to the annual data

```{r}
## This chunk for geting summary of ride_length variable only .
summary(annual_v2$ride_length)
```
```{r}
## This chunk for aggregate to variable and get ther summary .
aggregate(annual_v2$ride_length ~ annual_v2$member_casual, FUN = mean)
aggregate(annual_v2$ride_length ~ annual_v2$member_casual, FUN = median)
aggregate(annual_v2$ride_length ~ annual_v2$member_casual, FUN = max)
aggregate(annual_v2$ride_length ~ annual_v2$member_casual, FUN = min)
aggregate(annual_v2$ride_length ~ annual_v2$member_casual + annual_v2$day_of_week, FUN = mean)
```

## Create Visualization

```{r}
## This chunk to create a viz of rides number by rider type .
annual_v2 %>%
  mutate(weekday = wday(started_at, label = TRUE)) %>%
  group_by(member_casual, weekday) %>%
  summarise(number_of_rides = n(),average_duration = mean(ride_length)) %>%
  arrange(member_casual, weekday) %>%
  ggplot(mapping = aes(x = weekday, y = number_of_rides, fill = member_casual)) + geom_col(position = "dodge")
```
```{r}
## This chunk to create a viz for average duration .
annual_v2 %>%
  mutate(weekday = wday(started_at, label = TRUE)) %>%
  group_by(member_casual, weekday) %>%
  summarise(number_of_rides = n() ,average_duration = mean(ride_length)) %>%
  arrange(member_casual, weekday) %>%
  ggplot(mapping = aes(x = weekday, y = average_duration, fill = member_casual)) + geom_col(position = "dodge")
```
## Export Summary
```{r}
## This chunk for export the summary as csv file .
## Note : row.names parameter in False they not add a new column for names as 'X'
counts <- aggregate(annual_v2$ride_length ~ annual_v2$member_casual + annual_v2$day_of_week, FUN = mean)
write.csv(counts, "C:/Users/KasiolEr/Desktop/avg_ride_length.csv", row.names = FALSE)
```

